---
title: "Using CodeWAS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using CodeWAS}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(HadesExtras)
devtools::load_all(".")
#options("DEBUG_DATABASECONNECTOR_DBPLYR" = FALSE)
```

# Configuration 

Create a `CohortTableHandles` with a test cohort. 

```{r}
config_yaml <- "
  database: 
     databaseId: E1
     databaseName: Eunomia
     databaseDescription: Eunomia database
  connection:
    connectionDetailsSettings:
        dbms: eunomia
  cdm:
      cdmDatabaseSchema: main
      vocabularyDatabaseSchema: main
  cohortTable:
      cohortDatabaseSchema: main
      cohortTableName: test_cohort_table"

cat(config_yaml)
```

```{r}
config <- yaml::yaml.load(config_yaml)
```

```{r, eval=FALSE, echo=FALSE}
configurations <- yaml::read_yaml(testthat::test_path("config", "test_config.yml"))
configurationName <- "dev_bigquery"
configurationName <- "dev_eunomia"
config <- configurations[[configurationName]]
```

# Create CohortTableHandler

As in `CDMHandled`, to create a `CohortTableHandles`, for convenience `createCohortTableHandlerFromList` can be used. 
```{r}
cohortTableHandler <- createCohortTableHandlerFromList(config)
```

In addition to the checks performed by `CDMHandled`, `cohortTableHandler` includes a check on the creation of the `cohort` table. 

```{r}
cohortTableHandler$connectionStatusLog |> 
  reactable_connectionStatus()
```

# Import cohorts

```{r eval=FALSE}
cohortDefinitionSet <- CohortGenerator::getCohortDefinitionSet(
  settingsFileName = "testdata/asthma/Cohorts.csv",
  jsonFolder = "testdata/asthma/cohorts",
  sqlFolder = "testdata/asthma/sql/sql_server",
  cohortFileNameFormat = "%s",
  cohortFileNameValue = c("cohortId"),
  subsetJsonFolder = "testdata/asthma/cohort_subset_definitions/",
  packageName = "HadesExtras",
  verbose = FALSE
) |> tibble::as_tibble()

```


```{r}
cohortDefinitionSet <- CohortGenerator::getCohortDefinitionSet(
  settingsFileName = here::here("inst/testdata/asthma/Cohorts.csv"),
  jsonFolder = here::here("inst/testdata/asthma/cohorts"),
  sqlFolder = here::here("inst/testdata/asthma/sql/sql_server"),
  cohortFileNameFormat = "%s",
  cohortFileNameValue = c("cohortId"),
  subsetJsonFolder = here::here("inst/testdata/asthma/cohort_subset_definitions/"),
  #packageName = "HadesExtras",
  verbose = FALSE
) |> tibble::as_tibble()

```

Build cohort 
```{r}
cohortTableHandler$insertOrUpdateCohorts(cohortDefinitionSet)
```

```{r}
cohortTableHandler$getCohortsSummary() |> 
  rectable_cohortsSummary()
```


# CodWAS settings 

See list of analysis 

```{r}
analysis <- getListOfAnalysis()
analysis
```

```{r}
analysisIds  <- c(101, 141, 1, 2, 402, 702, 41, 301)

selectedAnalysis <- analysis  |> 
  dplyr::filter(analysisId %in% analysisIds)

selectedAnalysis
```



# Calculate codewas

Define covariates and time windows
```{r}
exportFolder <- file.path(tempdir(), "codeWAS")
dir.create(exportFolder, showWarnings = FALSE)

executeCodeWAS(
    exportFolder = exportFolder,
    cohortTableHandler = cohortTableHandler,
    cohortIdCases = 1,
    cohortIdControls = 2,
    analysisIds = analysisIds,
    covariatesIds = c(8507001, 1041),
    minCellCount = 1
    
)

```
  
  
Save results to sqlite
```{r}
csvFilesToSqlite(
  dataFolder = exportFolder,
  sqliteDbPath = file.path(tempdir(), "codeWAS.sqlite"),
  overwrite = TRUE, 
  analysis = "codeWAS"
)
```

connectTo sqlite
```{r}
DatabaseConnector::createConnectionDetails(
  dbms = "sqlite",
  server = file.path(tempdir(), "codeWAS.sqlite")
) |> 
  DatabaseConnector::connect() |> 
  DatabaseConnector::dbGetQuery("SELECT * FROM analysis_ref") 
```


# CohortDiagnostics

```{r}


# export results
tmp <- "/Users/javier/Downloads/tmp"

csvFilesToSqlite(
  dataFolder = exportFolder,
  sqliteDbPath = file.path(tmp, "codeWAS.sqlite"),
  overwrite = TRUE, 
  analysis = "codeWAS"
)

analysisSettings <- list(
  analysisType = "codeWAS",
  cohortIdCases =  1,
  cohortIdControls = 1,
  analysisIds = analysisIds,
  covariatesIds = 1002
)

yaml::write_yaml(analysisSettings, file.path(tmp, "analysisSettings.yaml"))

analysisResultsZipCsvPath <- file.path(tmp, "test_codeWAS.zip")
zip::zipr(zipfile = analysisResultsZipCsvPath, files = list.files(tmp, full.names = TRUE, recursive = TRUE))
```

